---
title: "Survival _II"
author: "Kelvin Nyongesa"
date: "2023-06-08"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r}
# Load the survival package
library(survival)

# Filter the dataset to include only observed events
kidney_observed <- kidney[kidney$status == 1, ]

# Create a survival object with the observed times
surv_obj <- Surv(kidney_observed$time)

# Fit the Kaplan-Meier estimator
km_fit <- survfit(surv_obj ~ 1)

# Access the estimated survival probabilities at each observed time point
km_survival <- km_fit$surv

# Print the estimated survival probabilities
library(knitr)
km_table<-summary(km_fit)
km_table
```

#############

```{r}
# Assuming you have a survfit object named 'survfit_obj'
summary_obj <- km_table

# Extract relevant information from the summary object
summary_data <- summary_obj$surv
lower_ci <- summary_obj$lower
upper_ci <- summary_obj$upper

# Create a data frame from the extracted information
summary_df <- data.frame(Survival = summary_data, Lower_CI = lower_ci, Upper_CI = upper_ci)

# Display the table using kable
kable(summary_df, format = "markdown", digits = 2, caption = "Survival Summary")


```

###########

```{r}
# Access the confidence interval values from the data set
ci_lower <- km_fit$lower 
ci_upper <- km_fit$upper

# Plot the survival function with 95% point wise confidence intervals
plot(km_fit, xlab = "Time", ylab = "Survival Function", main = "Kaplan-Meir Estimate of the Survival Function")
lines(km_fit, col = "blue")
lines(km_fit$time, ci_upper, type = "s", col = "red", lty = 2)
lines(km_fit$time, ci_lower, type = "s", col = "green", lty = 2)

# Add legend
legend("topright", legend = c("Survival Function", "upper","lower"),
       col = c("blue", "red","green"), lty = c(1, 2,2), bty = "n")
mtext("Figure 1: KM-Estimate", side = 1, line = 3, at = 0.5, cex = 1.2)
```

Estimating the Cummulative Hazard using the Nelson Aalen Estimator

#Create a survival object using the Surv() function, considering only the observed event times
```{r}
kidney_data<-kidney
surv_obj <- Surv(time = kidney_data$time[kidney_data$status == 1], event=kidney_data$status[kidney_data$status == 1])
```

#Apply the Nelson-Aalen estimator to estimate the cumulative hazard
```{r}
na_fit <- survfit(surv_obj ~ 1, data = kidney_data, type = "fh")
```

#Extract the cumulative hazard estimates and the confidence intervals
```{r}
na_cumhaz <- -log(na_fit$surv)
na_lower <- -log(na_fit$lower)
na_upper <- -log(na_fit$upper)
time <- na_fit$time
```

#Create a data frame with the cumulative hazard estimates and the confidence intervals
```{r}
library(knitr)
df_nelson <- data.frame(time, na_cumhaz, na_lower, na_upper)
na_table<-kable(df_nelson)
na_table
```

Plot of the Cumulative Hazard function
```{r}
# Plot the hazard function with 95% confidence intervals
plot(df_nelson$time, df_nelson$na_cumhaz, type = "s", col = "red",
     xlab = "Time", ylab = "Hazard", main = "Nelson-Aalen Estimate of the Cumulative Hazard")

# Add the upper and lower confidence intervals
lines(df_nelson$time, df_nelson$na_lower, type = "s", col = "purple", lty = 2)
lines(df_nelson$time, df_nelson$na_upper, type = "s", col = "green", lty = 2)
# Add a legend
legend("topleft", legend = c("Hazard Function", "Upper","Lower"),
       col = c("red","purple","green"), lty = c(1, 2,2), bty = "n")
# Add figure label
mtext("Figure 2: NA Estimator", side = 1, line = 3, at = 0.5, cex = 1.2)
```


